A Vocabulary to Describe RDF Constraints Generically

### 2015-05-25 paper/paper 2/rdf-constraints-vocabulary.tex
RDF Constraints Vocabulary

%
%
%\title{Formulating RDF Constraints and \\ Validating RDF Data Generically}
\title{Formulating RDF Constraints Generically and Validating RDF Data Out-Of-The-Box }
%\subtitle{}
%
\titlerunning{XXXXX}  % abbreviated title (for running head)
%                                     also used for the TOC unless
%                                     \toctitle is used
%
\author{Thomas Bosch\inst{1} \and Erman Acar\inst{2} \and Kai Eckert\inst{3}}
%
\authorrunning{XXXXX} % abbreviated author list (for running head)
%
%%%% list of authors for the TOC (use if author list has to be modified)
\institute{GESIS - Leibniz Institute for the Social Sciences, Germany\\
\email{thomas.bosch@gesis.org},\\ 
\and
University of Mannheim, Germany \\
\email{erman@informatik.uni-mannheim.de} 
\and
Stuttgart Media University, Germany \\
\email{eckert@hdm-stuttgart.de} 
}

@@

\begin{abstract}

More and more domains adopt Linked Data principles
and seek for ways to ensure data quality by validating RDF data according to predefined constraints.  In particular, as many communities are used to have this functionality in the XML world.
XML Schema is the standard language to formulate constraints to validate XML documents.
For RDF, however, there are several languages which can be used to express constraints on RDF data - but not a stand-alone standard one meeting the majority of RDF validation use cases. 
% Truth be told, we do may two things, blabla on and to blabla, however for the flow of the text we dont need to overkill the text. In abstract and introduction, we just give general idea so at such points we better focus on flow.  
We identified requirements to formulate constraints for  validating RDF data and mapped each requirement to a constraint type.
The majority of these constraint types can be expressed in description logics which provides their logical underpinning.

%In this paper, we provide a basic terminology and classification system for RDF constraints.\tb{The contributions of this paper are the following. (1)...}
%We developed a vocabulary to describe constraints of any RDF constraint type generically. 
%%(whether expressible in description logics or not).
%We show how to transform constraints, expressed by any constraint language \ms{$\alpha$}, into generically expressed constraints and into constraints represented by any other constraint language \ms{$\beta$} and discuss why these transformations are useful.
%We explain how to overcome the necessity to implement the validation of each constraint type for multiple constraint languages by providing mappings to generically expressed constraints.

The contributions of this paper are:
(1) we provide a basic terminology and classification system for RDF constraints,
(2) we develop a vocabulary to describe constraints of any constraint type generically and specified its underlying semantics,
(3) we show how to transform constraints which are expressed by any constraint language, into generically expressed constraints and discuss why these transformations are useful, and
(4) we explain how to overcome the problems to implement the validation of each constraint type for diverse constraint languages.

@@

\section{Introduction}
%The purpose of OWL 2 ontologies is to perform reasoning on RDF data and not to validate RDF data conforming to these ontologies.
%OWL 2 is based on the {\em non-unique name assumption} (nUNA) whereas RDF validation requires that different names represent different objects ({\em unique name assumption} (UNA)). 
%Reasoning in OWL 2 is based on the semantics of {\em open-world assumption} (OWA), i.e., a statement cannot be inferred to be false if it cannot be proved to be true  which fits its primary design purpose: to represent knowledge on the World Wide Web. 
%On the other hand, many RDF validation scenarios require the {\em closed-world assumption} (CWA) (i.e., a statement is inferred to be false if it cannot be proved to be true).
%This ambiguity in semantics is one of the main reasons why OWL 2 has not been adopted as a standard constraint language for RDF validation.  
%For this reason, we cannot use OWL 2 ontologies for RDF validation as we use XML Schemas, DTDs, RELAX NG, or Schematron to validate XML documents.

In 2013, the W3C organized the RDF Validation Workshop\footnote{\url{http://www.w3.org/2012/12/rdf-val/}}, 
where experts from industry, government, and academia discussed first use cases for RDF constraint\footnote{For simplicity reasons, we use the terms \emph{constraint types/constraints} instead of \emph{RDF constraint types/RDF constraints} in the remainder of the paper} formulation and RDF data validation.
In 2014, two working groups on RDF validation have been established to focus on RDF validation: 
the W3C RDF Data Shapes working group\footnote{\url{http://www.w3.org/2014/rds/charter}} and the DCMI RDF Application Profiles task group\footnote{\url{http://wiki.dublincore.org/index.php/RDF-Application-Profiles}}. 
Bosch and Eckert \cite{BoschEckert2014} collected the findings of these working groups and initiated a database of RDF validation requirements
which is available for contribution at \url{http://purl.org/net/rdf-validation}.
The intention is to collaboratively collect case studies, use cases, requirements, and solutions regarding RDF validation in a comprehensive and structured way. 
Bosch et al. identified 74 requirements to formulate constraints; each of them corresponding to a constraint type which can be expressed by at least one existing constraint language \cite{BoschNolleAcarEckert2015}.

Recently, RDF validation as a research field gained speed due to common needs of data practitioners. A typical example is the library domain that co-developed and adopted Linked Data principles very early. For libraries, the common description of resources are key business and they have a long tradition in developing and using interoperable data formats. While they embrace the openness of Linked Data and the data modeling principles provided by RDF, the data is still mostly represented in XML and this is unlikely to change any time soon. 
Among the reasons for the success of XML is the possibility to formulate fine-grained constraints to be met by the data and to validate the data according to these constraints using powerful systems like DTDs, XML Schemas, RELAX NG, or Schematron.
A typical example is the definition of a library record describing a book. There are clear rules which information has to be available to describe a book properly (required fields, like a title), but also how information like an ISBN number is properly represented. Libraries seek to make their own data reusable for general purposes, but also to enrich and interlink their own data. Checking if third-party data meets own requirements or validating existing data according to new needs for a Linked Data application are among common use cases for RDF validation.

\section{Motivation}

As there is no do-it-all language meeting the majority of requirements to formulate constraints and to validate RDF data,
%which can be used to express constraints and to validate RDF data, 
%(e.g. existential and universal quantification, cardinality restrictions, and exclusive or of properties)
language selection heavily depends on the individual use case.  
There are more than 10 candidate languages with different syntaxes and semantics, and we will focus on five which, in our opinion, are the most promising languages on being the standard constraint languages. Those are
\emph{Description Set Profiles (DSP)},
\emph{Resource Shapes (ReSh)}, 
\emph{Shape Expressions (ShEx)},
the \emph{SPARQL Inferencing Notation (SPIN)}, 
and the \emph{Web Ontology Language (OWL 2)}.
%\section{Ideas}
%constraint design patterns:
% - ontology design pattern.org
% - DSP uses other design pattern
% - constraint elements are the same / 
% - constraint / constraint elements / constraint design patterns / constraint language
%dependency between requirements / when this requirements is fulfilled then this is also fulfilled (e.g. min card and requ. car)
%min car more powerful than req. propery
The idea behind this paper is to achieve complete RDF validation requirements coverage by representing any constraint type (each of them corresponding to a requirement to formulate constraints on and to validate RDF data) in a generic way using a lightweight vocabulary which consists of only a few terms (see Section \ref{sec:vocabulary}).

\emph{Cardinality restrictions} can be expressed either generically (\emph{generic constraints}) by \emph{description logics (DL)} 
%and extensions  
or specifically (\emph{specific constraints}) by domain-specific constraint languages such as \emph{OWL 2}, \emph{DSP}, \emph{ShEx}, \emph{ReSh}, and \emph{SPIN}.
The knowledge representation formalism {\em DL}, with its  well-studied theoretical properties, provides the foundational basis to express constraints in a generic way. 
For that reason, we map constraint types to \emph{DL} statements.

With \emph{minimum qualified cardinality restrictions} (\emph{R-75/81})\footnote{Requirements/constraint types are uniquely identified by alphanumeric technical identifiers like \emph{R-75-MINIMUM-QUALIFIED-CARDINALITY-ON-PROPERTIES}}, researchers from the library domain may restrict
that \emph{publications} must have at least one \emph{author} which must be a \emph{person}.
This constraint can be represented generically (\emph{generic constraint}) in \emph{DL} {\small\ms{(Publication $\sqsubseteq$ $\geq$1 author.Person)}}, but also specifically (\emph{specific constraint}) by multiple constraint languages:

\begin{ex}
# OWL 2:
Publication
    a owl:Restriction ;
    owl:minQualifiedCardinality 1 ;
    owl:onProperty author ;
    owl:onClass Person .

# ShEx:
Publication { author @Person{1, } }

# ReSh:
Publication a rs:ResourceShape ; rs:property [
    rs:name "author" ; rs:propertyDefinition author ;
    rs:valueShape Person ;
    rs:occurs rs:One-or-many ; ] .

# DSP:
[   a dsp:DescriptionTemplate ; 
    dsp:resourceClass Publication ; 
    dsp:statementTemplate [ a dsp:NonLiteralStatementTemplate ;
        dsp:minOccur 1 ; dsp:maxOccur "infinity" ; 
        dsp:property author ; 
        dsp:nonLiteralConstraint [ a dsp:NonLiteralConstraint ;
            dsp:valueClass Person ] ] ] .
\end{ex}

As every \emph{publication} must have at least one \emph{author} which must be a \emph{person}, 
and since the book \emph{The-Lord-Of-The-Rings} is a \emph{publication} {\small\ms{(rdf:type(The-Lord-Of-The-Rings, Publication)),}} 
it must have at least one \emph{author} relationship to a \emph{person}.
Constraint violations are raised either (1) if
\emph{The-Lord-Of-The-Rings} does not have any \emph{author} relationship, or (2) if 
it has an \emph{author} which is not a \emph{person} \\
{\small\ms{(author(The-Lord-Of-The-Rings, Tolkien)}, \ms{rdf:type(Tolkien, Hobbit)),}} 
or (3) if it has an \emph{author} for which no class is assigned
{\small\ms{(author(The-Lord-Of-The-Rings, Tolkien)).}}
In contrast, \emph{The-Lord-Of-The-Rings} is a valid \emph{publication}, if it is connected to a \emph{person} via the property \emph{author}
{\small\ms{(author(The-Lord-Of-The-Rings, Tolkien)}, \ms{rdf:type(Tolkien, Person)).}}

As there is no standard way to formulate constraints, 
semantically equivalent \emph{cardinality restrictions} may be represented by multiple constraint languages that have different syntax and semantics.
This causes confusion and weakens the common understanding between several parties about the semantics of particular constraints. Also, it raises  questions on how to ensure high data quality. 
Therefore, when choosing a constraint language, say \ms{$\alpha$}, to express a constraint of an arbitrary constraint type, it should be possible to transform this constraint into a semantically equivalent constraint expressed by another constraint language \ms{$\beta$} (see Section \ref{sec:transformations}). 
This is important in order to enhance the interoperability of constraint languages and to resolve misunderstandings and ambiguities in the communication of RDF data producers and consumers, 
since constraint transformations avoid the necessity to understand several constraint languages.

We provide a validation environment\footnote{Available \url{http://purl.org/net/rdfval-demo}} which enables to validate RDF data according to constraints expressed by diverse  constraint languages.
As a second step, the \emph{RDF Validator} is extended by validating RDF data conforming to constraints of any constraint type represented by any constraint language,  
to perform validation on semantically equivalent \emph{specific constraints}, expressed by different constraint languages, 
i.e., validation is independent from the used constraint language (see Section \ref{sec:validation}). 
%i.e., validation leads to identical validation results - independently from the used constraint language (see section \ref{sec:validation}).
We are able to offer an already implemented validation mechanism for any \emph{specific constraint} out-of-the-box, 
(1) if we implement the validation of the corresponding \emph{generic constraint type} once and 
(2) if we transform semantically equivalent \emph{specific constraints} into generically expressed constraints.

%When domain experts use graphical user interfaces to define constraints in a user-friendly way (without the necessity to express constraints on their own), 
%constraints can be mapped to either a specific or a generic constraint, as for both the validation is provided in the background.    
%So, the user does not have to know how to express specific constraints using any constraint language.

%Existing constraint languages can be extended and new constraint languages can be developed, if existing constraint languages are not sufficient to represent needed constraint types.

%If (1) a new constraint type is expressed generically, (2) the \emph{generic constraint type} is mapped to \emph{specific constraint types} (expressed by constraint languages), and (3) the validation is implemented for this particular \emph{generic constraint type},
%the validation of \emph{specific constraint types} is provided out of the box.
%As a consequence, constraint language designers do not have to implement the validation of their constraint languages and each semantically equivalent \emph{specific constraint type} is validated the same way (i.e. leading to identical validation results).

%This paper aims to address two main \textbf{audiences}:
%(1) RDF data providers and consumers seeking how to ensure high data quality and thus for ways to enhance the interoperability and common understanding of constraints expressible by multiple constraint languages and  
%(2) RDF practitioners thinking of how to provide aligned implementations for the validation of both existing an newly developed constraint languages.
The main contributions of this paper are:
(1) we provide a basic terminology and classification system for RDF constraints (Section \ref{sec:vocabulary}),
(2) we developed a vocabulary to describe constraints of any RDF constraint type generically and specified its underlying semantics (Section \ref{sec:vocabulary}),
(3) we show how to transform constraints, expressed by any constraint language \ms{$\alpha$}, into generically expressed constraints and into constraints represented by any other constraint language \ms{$\beta$} and discuss why these transformations are useful (Section \ref{sec:transformations}), and
(4) we explain how to overcome the necessity to implement the validation of each constraint type for multiple constraint languages (Section \ref{sec:validation}).
%(if the validation is implemented once for the corresponding \emph{generic constraint type} to which the \emph{specific constraint type} is mapped)

%The remainder of the paper is as follows.

%\begin{itemize}
	%\item Any specific constraint (expressed by any constraint language) can be transformed into a generic constraint (expressed by the generic constraint language)
	%\item Any specific constraint (expressed by any constraint language A) can be transformed into any specific constraint (expressed by any constraint language B)
	%\item Any specific constraint (expressed by any constraint language) can be validated automatically (if an automatic validation is defined only once for corresponding generic constraint)
	%\item We define a basic terminology and classification system for RDF constraints
	%\item We developed an ontology describing any RDF constraint generically (whether expressible in DL or not)%which can also be expressed by any constraint language
	%\item We present that any RDF constraint can be described by the developed ontology - whether the constraint can be expressed in DL or not 
	%\item We show how to transform specific constraints (expressed by any constraint language \ms{$\alpha$}) into generic constraints and into specific constraints (expressed by any other constraint language \ms{$\beta$})
  %\item We explain that any specific constraint can be validated automatically (if the validation is implemented only once for the corresponding generic constraint)
%\end{itemize}

%The \textbf{benefits} of our framework are:
%\begin{itemize}
	%\item Any specific constraint (expressed by any constraint language) can be transformed into a generic constraint (expressed by the generic constraint language)
	%\item Any specific constraint (expressed by any constraint language A) can be transformed into any specific constraint (expressed by any constraint language B)
	%\item Any specific constraint (expressed by any constraint language) can be validated automatically (if an automatic validation is defined only once for corresponding generic constraint)
%\end{itemize}

%\section{Running Example}
%
%Table \ref{tab:cardinality-restriction-mapped-to-vocabulary} shows the mapping between the \emph{cardinality restriction} represented in DL 
%and the \emph{cardinality restriction} expressed by our developed vocabulary to describe any RDF constraint generically (see section~\ref{sec:ontology}):
%
%\begin{table}
  %\scriptsize
  %\sffamily
  %\vspace{0cm}
	%\centering
		%\begin{tabular}{l|l|l|l|l|l|l}
      %\textbf{c. type} & \textbf{context class} & \textbf{left p. list} & \textbf{right p. list} & \textbf{classes} & \textbf{c. element} & \textbf{c. value} \\
      %\hline
      %property & Publication & author & - & Person & $\geq$ & 1 \\
		%\end{tabular}
	%\caption{Cardinality Restriction Expressed Generically}
	%\label{tab:cardinality-restriction-mapped-to-vocabulary}
%\end{table}

%According to this vocabulary, each constraint is either a constraint on classes or a constraint on properties.
%The generically expressed cardinality restriction is represented in RDF as follows:
%
%\begin{ex}
%[   a PropertyConstraint ;
    %contextClass Publication ;
    %leftProperties ( author ) ;
    %classes ( Person ) ;
    %constrainingElement ">=" ;
    %constrainingValue 1 ] .
%\end{ex}

%Bosch and Eckert\cite{BoschEckert2014-2} use SPIN as basis to define a
%validation environment in which the validation of any constraint language\footnote{The only limitation is that constraint languages must be represented in RDF} can be implemented by representing them in SPARQL. 
%The SPIN engine checks for each resource if it satisfies all constraints (associated with its assigned classes) and generates a result RDF graph containing information about all constraint violations.
%The generically expressed \emph{cardinality restriction} above is validated within the validation environment (see section \ref{sec:validation}) by means of a SPARQL CONSTRUCT query.

%\begin{ex}
%[   a PropertyConstraint ;
    %contextClass ?cc ;
    %leftProperties ( ?p1 ) ;
    %classes ( ?c1 ) ;
    %constrainingElement ">=" ;
    %constrainingValue ?cv ] .
%?subject rdf:type ?cc .
%BIND ( qualifiedCardinality( ?subject, ?p1, ?c1 ) AS ?c ) .
%FILTER ( ?c < ?cv ) .		  
%\end{ex}

@@

Constraints can be expressed either generically (\emph{generic constraints}) by the \emph{RDF-CV}   
or specifically (\emph{specific constraints}) by domain-specific constraint languages.
As the \emph{RDF-CV} describes constraints generically, the vocabulary does not distinguish constraints according to the dimension \emph{universality}. 
The majority of the constraint types can be expressed in \emph{DL}.
In contrast, there are certain constraint types which cannot be expressed in \emph{DL}, but are also expressible by the \emph{RDF-CV} . 
\emph{Complex constraints} encompass \emph{simple constraints} (atomic constraints) and/or further \emph{complex constraints},
i.e. \emph{DL} statements representing \emph{complex constraints} are created out of \emph{DL} statements standing for atomic and composed constraints (if expressible in \emph{DL}). 
Simple constraints may be applied to either properties (\emph{property constraints}) or classes (\emph{class constraints}).
The \emph{RDF-CV} does not contain any terms standing for \emph{simple} and \emph{complex constraints}, since \emph{context classes} of \emph{simple constraints} are just reused within \emph{complex constraints} (\emph{simple constraints} associated with \emph{context classes} must hold for individuals of these \emph{context classes}). 
%\tb{Erman: If you say its expressible in RDF-CV, than you impose that RDF-CV is a language rather than a vocabulary. "in different contexts they call it signature which is the same thing". Generally, a language uses vocabulary to construct sentences, and the set of those sentences, is the language.} (see section \ref{sec:evaluation})
As a consequence, the distinction of \emph{property} and \emph{class constraints} is sufficient to describe constraints of all possible constraint types.
\er{use RDF-CV / not: expressible in RDF-CV / RDF-CV is a vocabulary, not a language}

@@

\section{Related Work}

\tb{0.75 page / explain how typical constraints are expressed by different constraint languages}

>>>>

%
%
%\title{Formulating RDF Constraints and \\ Validating RDF Data Generically}
\title{A Vocabulary to Describe RDF Constraints Generically}
%\subtitle{}
%
\titlerunning{XXXXX}  % abbreviated title (for running head)
%                                     also used for the TOC unless
%                                     \toctitle is used
%
\author{Thomas Bosch\inst{1} \and Kai Eckert\inst{3}}
%
\authorrunning{XXXXX} % abbreviated author list (for running head)
%
%%%% list of authors for the TOC (use if author list has to be modified)
\institute{GESIS - Leibniz Institute for the Social Sciences, Germany\\
\email{thomas.bosch@gesis.org},\\ 
\and
Stuttgart Media University, Germany \\
\email{eckert@hdm-stuttgart.de} 
}

@@

\begin{abstract}
...

@@

@@

Constraints can be expressed either generically (\emph{generic constraints}) by the \emph{RDF-CV}   
or specifically (\emph{specific constraints}) by domain-specific constraint languages.
As the \emph{RDF-CV} describes constraints generically, the vocabulary does not distinguish constraints according to the dimension \emph{universality}. 
The majority of the constraint types can be expressed in \emph{DL}.
In contrast, there are certain constraint types which cannot be expressed in \emph{DL}, but are also expressible by the \emph{RDF-CV} . 
\emph{Complex constraints} encompass \emph{simple constraints} (atomic constraints) and/or further \emph{complex constraints},
i.e. \emph{DL} statements representing \emph{complex constraints} are created out of \emph{DL} statements standing for atomic and composed constraints (if expressible in \emph{DL}). 
Simple constraints may be applied to either properties (\emph{property constraints}) or classes (\emph{class constraints}).
The \emph{RDF-CV} does not contain any terms standing for \emph{simple} and \emph{complex constraints}, since \emph{context classes} of \emph{simple constraints} are just reused within \emph{complex constraints} (\emph{simple constraints} associated with \emph{context classes} must hold for individuals of these \emph{context classes}). 
%\tb{Erman: If you say its expressible in RDF-CV, than you impose that RDF-CV is a language rather than a vocabulary. "in different contexts they call it signature which is the same thing". Generally, a language uses vocabulary to construct sentences, and the set of those sentences, is the language.} (see section \ref{sec:evaluation})
As a consequence, the distinction of \emph{property} and \emph{class constraints} is sufficient to describe constraints of all possible constraint types.

@@
